\documentclass[titlepage]{scrartcl}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{hyperref}
\usepackage{mathtools}
\usepackage{pythonhighlight}
\usepackage{enumitem}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{algorithm}
\usepackage[maxnames=4]{biblatex}
\usepackage[outputdir=out, cache=true]{minted}


\usepackage{tikz}
\usetikzlibrary{chains,shapes.multipart}
\usetikzlibrary{shapes,calc}
\usetikzlibrary{automata,positioning}
\usetikzlibrary{arrows, decorations.markings}
\usetikzlibrary{shapes}

\usepackage{pgfplots}
\pgfplotsset{compat=1.16}
\usepgfplotslibrary{statistics} % For boxplots 

\addbibresource{literature.bib}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\newtheorem{remark}[theorem]{Remark}

\makeatletter
\newcounter{classdef}
\newenvironment{classdef}[1][htb]
  {\floatname{algorithm}{Data Structure}
   \let\c@algorithm\c@classdef
   \begin{algorithm}[#1]%
  }{\end{algorithm}}
\makeatother

\BeforeBeginEnvironment{minted}{\vspace{0.3em}}
\AfterEndEnvironment{minted}{\vspace{-0.6em}}

\author{Michael Markl}
\title{Computation of Dynamic~Prediction~Equilibria}
\subtitle{A software project as part of the M.Sc. Mathematics at~the~University~of~Augsburg.}
\publishers{Supervised by Prof. Dr. Tobias Harks}

\newcommand{\R}{\mathbb{R}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\diff}{\,\textnormal{d}}
\newcommand{\contFcts}[2]{{C}{\left({#1},{#2}\right)}}
\newcommand{\emptyArg}{\,\boldsymbol{\cdot}\,}
\newcommand{\abs}[1]{\left| #1 \right|}

\newcommand{\minv}[1]{#1^{\leftarrow}}
\newcommand{\rev}[1]{#1^{\leftarrow}}
\newcommand{\id}{\mathbf{id}}
\newcommand{\bigO}{\mathcal{O}}
\newcommand{\firstN}[1]{[#1]}
\newcommand{\restrict}[2]{#1\vert_{#2}}
\newcommand{\colonequiv}{\vcentcolon\mspace{-1.2mu}\equiv}
\newcommand{\equivcolon}{\equiv\mspace{-1.2mu}\vcentcolon}
\newcommand{\restrictUpTo}[2]{#1_{\leq #2}}


% Flow-specifics
\newcommand{\capa}{\nu}
\newcommand{\transit}{\tau}
\newcommand{\outEdges}[1]{\delta^-_{#1}}
\newcommand{\inEdges}[1]{\delta^+_{#1}}
\newcommand{\infl}[1][f]{{#1}^+}
\newcommand{\outfl}[1][f]{{#1}^-}
\newcommand{\bal}[2]{b_{#1,#2}}
\newcommand{\qulen}{q}
\newcommand{\predq}{\hat q}
\newcommand{\predc}{\hat c}
\newcommand{\predT}{\hat T}
\newcommand{\predl}{\hat l}
\newcommand{\predE}{\hat E}
\newcommand{\paths}{\mathcal{P}}
\newcommand{\code}[1]{\mintinline{python}{#1}}
\newcommand{\depl}{\textnormal{depl}}
\newcommand{\QueueFct}{\mathcal{Q}}

\newcommand{\predZ}{{\text{Z}}}
\newcommand{\predC}{{\text{C}}}
\newcommand{\predL}{{\text{L}}}
\newcommand{\predRL}{{\text{RL}}}
\newcommand{\predML}{{\text{ML}}}
\newcommand{\predP}{{\text{P}}}


\begin{document}
    \maketitle

    \clearpage
    \tableofcontents

    \clearpage
    \section{Introduction}

    When analyzing real-world traffic or communication networks, often simulations are used to retrieve structural information on the performance of the network given certain demands.
    Starting from a more theoretical perspective, equilibria of underlying game-theoretic models propose to give more fundamental insights.
    In this software project, we aim to compute such an equilibrium called Dynamic Prediction Equilibrium (DPE) which was recently introduced in~\cite{mainpaper}.
    The fundamental dynamics adhere to Vickrey's fluid queuing model in which infinitesimal particles enter edges.
    Once the amount of flow exceeds an edge's capacity, a point queue will form in front of it.

    In a DPE, each commodity has a source-sink pair as well as some mean of predicting future travel times of any edge.
    An agent of a commodity, an infinitesimal particle traveling from the commodity's source $s$ to its sink $t$, computes a shortest path to the sink $t$ every time it arrives at a node $v$ by instantiating a new prediction on the future edge costs. Then, it enters an outgoing edge of $v$ lying on that path.
    The predictions of a commodity $i$ are modeled as a function $\predq_{i,e}(\theta, \bar\theta, q)$ that given the past queue lengths $q: \R\rightarrow \R_{\geq0}^E$ will return the predicted queue length of edge $e$ at time $\theta$ as predicted at time $\bar\theta$.
    If all commodities use a perfect predictor, which predicts all queue lengths correctly, the resulting equilibrium flow is called a Dynamic Equilibrium which has been studied thoroughly in literature, for example in~\cite{Cominetti2015}.
    

    In this document, we first take a closer look at general time-dependent cost functions following the First-In-First-Out (FIFO) rule in Section~\ref{sec:fifo-costs}:
    We discuss a duality of arrival and departure times that can be used for a simple Dijkstra-based algorithm to compute all so-called \emph{active edges}.
    In Section~\ref{sec:equilibrium-flows} we describe the mathematical flow model and define DPEs within this model.
    The computation of an approximation of such an equilibrium is described in detail in Section~\ref{sec:compute-dpes}.
    Finally, we run our developed algorithms in Section~\ref{sec:results} on a small synthetic network and multiple real world traffic networks.

    All code snippets presented in this work use the Python 3.8 syntax where type hints (e.g. \code{i: int}) are often used to make the underlying data encoding clearer.
    The presented source code is often a simplified version of the original code such that the reader is not distracted by technical details which might improve the performance.
    Moreover, in this work we abstract from any kind of rounding errors due to the fixed size floating point numbers.
    It is worth mentioning though, that fixing these kinds of errors was quite laborious. 
    The full source code of this software project is publicly available at \url{https://github.com/schedulaar/predicted-dynamic-flows}.

    \clearpage
    \input{fifo-cost-functions.tex}

    \clearpage
    \section{Equilibrium Flows}\label{sec:equilibrium-flows}

    The goal of this project is to compute a specific kind of equilibrium flows.
    This section gives the appropriate introduction in the mathematical model of the underlying flows and introduces the so-called Dynamic Prediction Equilibria that we aim to compute.
    
    \input{dynamic-flows.tex}
    \input{dynamic-prediction-equilibrium.tex}

    \clearpage
    \input{flow-computation.tex}
    
    \clearpage
    \input{computational-results.tex}

    \clearpage
    \section{Conclusion}

    In this report we have introduced a method to compute $\varepsilon$-DPEs.
    We introduced several predictors and their implementations and showed how machine learning methods can be embedded in this simulation.
    This approach can be used to analyze the use of machine learning based navigation systems in everyday traffic.
    The benefit over other traffic simulations is that it follows a very concise and clear mathematical model that can also be analyzed theoretically.

    We want to conclude this report with several aspects for future work.

    First, the current version of the software does not come with a general-graph visualization of the calculated flows.
    It is clear, that a good visualization can improve the understanding and intuition of the underlying model.
    Moreover, it is interesting to see which edges become bottlenecks for a given scenario and network and inspect how changes to these edges or the insertion of edges influences the resulting flow.

    Second, the machine learning model used here can still be improved as can be seen in Section~\ref{sec:results}.
    It might be worth changing the feature set given to the learning algorithm:
    The queue length of surrounding edges might not be the only indicator of an upcoming queue at the future.
    For example if a surrounding edge has a very high capacity and transports an amount of flow close to its capacity, this flow cannot be detected in the current feature set, because no queue forms in front of the edge. 
    However, changing the feature set to include more than the past queue lenghts also changes the mathematical model of the predictor functions $\predq_{i,e}(\emptyArg, \emptyArg, \emptyArg)$.
    Changing the arguments of this function might imply that the existence of equilibria as given by Theorem~\ref{thm:dpe-existence} does not hold anymore.
    
    Finally, the quality of approximation of $\varepsilon$-DPEs is still to be analyzed.
    A desirable result would be that for $\varepsilon$ tending to $0$, the $\varepsilon$-DPE converges to a real DPE.
    However, in this report we leave this question open for further research.

    \clearpage
    \printbibliography
\end{document}
